paper1-INTERACT-analysis

### 2024-09-09 paper1-INTERACT-analysis.Rmd
updated figures to read "phase" instead of "wave", to match with new language in the paper. Also reviewed the code to ensure that models were adjusted for the correct variables as I was revising the proofs. I noticed that the fully adjusted model for objective one was missing education level, so I added that in and re-ran the model. It was also missing from the first of the three interaction plots. Adding this in did not meaninfully change the results -- only the CI shifted by 0.01. Similarly, I noticed that income was missing from the code for the fully adjusted model for objective 2. I added

*sd_ppl: Person-level dataset*: used in version 1 of Table 1 (with person-level characteristics)
```{r}
  sd_ppl<-sd_pa %>% ungroup() %>% 
                    dplyr::select(interact_id,income,gender,age,age_cat,children,
                           education,employment,wave_id,city_id,both)
  sd_ppl<-distinct(sd_ppl)
  table(sd_ppl$wave_id,sd_ppl$city_id) # one participant from Saskatoon in wave 1 has gone missing?
```

@@

@@

Plot of stringency index over time (in black and white)
```{r,echo=FALSE}
  si_plot<-ggplot(si[si$province %in% c("British Columbia","Saskatchewan","Quebec") & 
                      si$date <= "2021-02-28",],aes(x=date,y=stringencyIndex*10,color=province)) +
           geom_rect(aes(xmin=ymd('2020-09-10'),
                  xmax = ymd('2021-02-28'),
                  ymin = -Inf,
                  ymax = Inf), fill = 'grey95', color='grey95',alpha = 0.75) +
           geom_line(size=0.6) + 
           theme_classic() + 
           scale_colour_grey(labels=c("British Columbia\n(Vancouver)\n",
                                      "Quebec\n(Montreal)\n",
                                      "Saskatchewan\n(Saskatoon)\n")) +
           labs(#title="Stringency of COVID-19 restrictions over time",
                 x="Date",y="Stringency index") +
           guides(color=guide_legend(title="Province (City)\n")) +
           theme(legend.title=element_text(face="bold",size=9),
                 legend.text = element_text(size=8),
                 axis.title = element_text(face="bold",size=9),
                 axis.text = element_text(size=8),
                 plot.title = element_text(face="bold",size=10,hjust=0.5)) +
           annotate(geom="text", x=as.Date('2020-12-04'), y=10, label="Wave 2 data collection period",
                    color="grey30",size=2.75)

@@

  ggsave(si_plot,file="figures//figure4.png",width=5.5,height=3.25)
  ggsave(si_plot,file="submission to IJE//figure4.png",width=175,height=95,unit="mm") # for IJBNPA
```

@@

**Figure 1**: Distribution of daily minutes of wear time and moderate-to-vigorous physical activity among valid person-days, by wave
```{r,echo=FALSE}
wave.labs<-c("Wave 1","Wave 2")
names(wave.labs)<-c("wave 1","wave 2")

  wear_plot<-ggplot(sd_valid,aes(x=wear)) + 
    geom_histogram(bins=50,aes(fill=wave_id,color=wave_id)) +
    theme_classic() +
    labs(x="Minutes",
         y="Frequency",
         title="Wear time" #,
         #subtitle="Wave 1 (dark blue) vs. Wave 2 (light blue)"
         ) +
    geom_vline(xintercept=600,color="black",linetype="dashed") +
    facet_grid(vars(wave_id),
               labeller = labeller(wave_id=wave.labs)) +
    # scale_fill_manual(values=c("wave 1"="#0a9ffe", # green = "#81DA66"
    #                             "wave 2"="#fe7701"), # darker orange #fe7701 #ff8c09
    #                    labels=c("Wave 1","Wave 2")) + 
    scale_fill_grey(labels=c("Wave 1","Wave 2")) +
    scale_color_grey(labels=c("Wave 1","Wave 2"))+
    theme(axis.title.x = element_text(face="bold",
                                       size=10),
          axis.text.x = element_text(angle=90,vjust = 0.5, hjust=1,size=8),
        # axis.title.y = element_text(#face="bold",
        #                              size=10),
          axis.title.y = element_blank(),
          axis.text.y = element_text(size=8),
          plot.title = element_text(face=2,
                                    hjust=0.5,
                                    size=10),
          legend.position="none",
          strip.text.y = element_text(face=2))

  mvpa_plot<-ggplot(sd_valid,aes(x=mean_mvpa_sd)) + 
    geom_histogram(bins=50,aes(fill=wave_id,color=wave_id)) +
    theme_classic() +
    labs(x="Minutes",
         y="Frequency",
         title="Moderate-to-vigorous physical activity" #,
         #subtitle="Wave 1 (dark blue) vs. Wave 2 (light blue)"
         ) +
    facet_grid(vars(wave_id),
               labeller = labeller(wave_id=wave.labs)) +
    # scale_fill_manual(values=c("wave 1"="#0a9ffe", # green = "#81DA66"
    #                             "wave 2"="#fe7701"), # darker orange #fe7701 #ff8c09
    #                    labels=c("Wave 1","Wave 2")) + 
    scale_fill_grey(labels=c("Wave 1","Wave 2")) +
    scale_color_grey(labels=c("Wave 1","Wave 2"))+
    theme(axis.title.x = element_text(face="bold",
                                       size=10),
          axis.text.x = element_text(angle=90,vjust = 0.5, hjust=1,size=8),
        # axis.title.y = element_text(#face="bold",
        #                              size=10),
          axis.title.y = element_blank(),
          axis.text.y = element_text(size=8),
          plot.title = element_text(face=2,
                                    hjust=0.5,
                                    size=10),
          legend.position="none",
          strip.text.y = element_text(face=2))

  # titleMain=ggpubr::text_grob("Distribution of daily minutes of wear time and moderate-to-vigorous physical activity:\nWave 1 (2018-2019) vs. Wave 2 (2020-2021)\n",
  #                             size =10,face="bold")  
  titleY=ggpubr::text_grob("Frequency",
                            size =10,rot=90,face="bold")

  figure1<-grid.arrange(wear_plot,mvpa_plot,ncol=2,
               #top=titleMain, # removing title for IJBNPA (titles should be in the text, not in the figure)
               left=titleY)

  ggsave(figure1,file="figures//figure1.png",width=6,height=4)
  ggsave(figure1,file="submission to IJE//figure1.png",width=175,height=100,unit="mm")
```

**Figure 2:** Trend in daily minutes of moderate-to-vigorous physical activity for each wave
```{r,echo=FALSE,warning=FALSE}
  sd_valid$doy3<-sd_valid$date
  sd_valid$doy3<-gsub("2018|2020","2000",sd_valid$doy3)
  sd_valid$doy3<-gsub("2019|2021","2001",sd_valid$doy3)
  sd_valid$doy3<-as.Date(sd_valid$doy3)

  figure2<-ggplot(sd_valid[sd_valid$wave_id=="wave 1",],aes(x=doy3,y=mean_mvpa_sd,color=wave_id)) +
                          geom_point(alpha=0.15,size=1)+
                          geom_point(data=sd_valid[sd_valid$wave_id=="wave 2",],aes(x=doy3,y=mean_mvpa_sd),
                                     alpha=0.15,size=1) +
                          scale_colour_manual(values=c("wave 1"="grey65",
                                                       "wave 2"="black"),
                                              labels=c("Wave 1 (2018-2019)","Wave 2 (2020-2021)")) +
                          geom_smooth(method=lm)+
                          geom_smooth(data=sd_valid[sd_valid$wave_id=="wave 2",],aes(x=doy3,y=mean_mvpa_sd),method=lm) +
                          theme_classic()+
                          labs(#title = "Daily minutes of moderate-to-vigorous physical activity by day of the year: ", # removing title for IJBNPA
                               #subtitle = "Wave 1 (2018-2019) vs. Wave 2 (2020-2021)",
                               y = "Daily minutes of moderate-to-vigorous physical activity",
                               x = "Date")+
                          guides(color=guide_legend("")) +
                          theme(axis.title = element_text(face="bold",size=10),
                                axis.text.x = element_text(angle=90,vjust = 0.5, hjust=1,size=8),
                                axis.text.y = element_text(size=8),
                                plot.title = element_text(face="bold",hjust=0.5,size=10),
                                plot.subtitle = element_text(hjust=0.5,size=10),
                                legend.position="bottom",
                                legend.key.size = unit(1,"line"),
                                legend.spacing.x = unit(0.25, 'cm'),
                                legend.margin=margin(-3, 0, 0, 0),
                                legend.text = element_text(size=8,margin = margin(r = 0.4, unit = 'cm'))) 
  figure2
  ggsave(figure2,file="figure2.png",width=6,height=4)
  ggsave(figure2,file="submission to IJE//figure2.png",width=175,height=120,unit="mm")
```

@@

*Model: Adjusting for city, age, employment status, income, temp, precipitation and wear time*
```{r,warning=FALSE}
  wave_covar_valid_fully<-glmer.nb(mean_mvpa_sd~ wave_id + city_id + wearRS +
                                                 age_cat + employment + income +
                                                 mean_temp + total_precip + 
                                                 (1|interact_id),
                          data=sd_valid[sd_valid$age_cat!="missing",])
```

@@

**Model**
```{r,warning=FALSE}
  both_valid_adj_wear<-glmer.nb(mean_mvpa_sd~ wave_id + 
                                              gender + age_cat + income +
                                              mean_temp + total_precip + wear +
                                       (1|interact_id),
                          data=sd_valid_both) 
```

@@

#### Plot effect modifiers of interest (Figure 3): Income, employment status and education

@@

# Add a income-wave interaction term to the adjusted model  
income_covar_valid_final<-glmer.nb(mean_mvpa_sd~ wave_id + city_id + wave_id*income + wearRS +
                                                age_cat + employment + income +
                                                mean_temp + total_precip + 
                                                (1|interact_id),
                          data=sd_valid[sd_valid$age_cat!="missing",])

@@

# create factor variables for both wave_id and income             
income_Inter.HandPick$Income <- factor(income_Inter.HandPick$income,
                      levels=c("<$50,000","$50,000 to $99,999","$100,000 or more","don't know/prefer not to answer"
                               ),
                      labels=c("<$50,000","$50,000-$99,999","$100,000+","No answer"
                               ))
income_Inter.HandPick$wave <- factor(income_Inter.HandPick$wave_id,
              levels=c("wave 1","wave 2"),
              labels=c("Wave 1", "Wave 2"))

@@

# create factor variables for both wave_id and education level          
education_Inter.HandPick$Education <- factor(education_Inter.HandPick$education,
                      levels=c("trade school or college","secondary school","university or graduate degree"),
                      labels=c("College/trade school","High school","University/grad degree"))
education_Inter.HandPick$wave <- factor(education_Inter.HandPick$wave_id,
              levels=c("wave 1","wave 2"),
              labels=c("Wave 1", "Wave 2"))

@@

# create factor variables for both wave_id and employment status             
employment_Inter.HandPick$Employment <- factor(employment_Inter.HandPick$employment,
                      levels=c("employed","unemployed","retired and not working","other"),
                      labels=c("Employed","Unemployed","Retired","Other"))
employment_Inter.HandPick$wave <- factor(employment_Inter.HandPick$wave_id,
              levels=c("wave 1","wave 2"),
              labels=c("Wave 1", "Wave 2"))

@@

**Create Figure 3**
```{r,echo=FALSE,warning=FALSE}
titleY=ggpubr::text_grob("Daily minutes of moderate-to-vigorous\nphysical activity",size =12,rot=90,face="bold")

em_figures<-grid.arrange(income_Plot.HandPick,
                         employment_Plot.HandPick,
                         education_Plot.HandPick,
            left=titleY,
            widths=c(1,1,1),
            heights=c(1),
            nrow=1)
ggsave(em_figures,file="figures//figure3.png",width=175,height=85,units="mm")
ggsave(em_figures,file="submission to IJE//figure3.png",width=175,height=85,units="mm")
```

@@

**Model**
```{r,warning=FALSE}
  si_all_cov_fully<-glmer.nb(mean_mvpa_sd~ stringencyIndex + wear + city_id + 
                          mean_temp + total_precip + 
                          age_cat + 
                          gender + 
                          children + 
                          employment + 
                          (1|interact_id),
                          data=sd_valid[sd_valid$wave_id=="wave 2",])
```

>>>>

*sd_ppl: Person-level dataset*: used in version 1 of Table 1 (with person-level characteristics)
```{r}
  sd_ppl<-sd_pa %>% ungroup() %>% 
                    dplyr::select(interact_id,income,gender,age,age_cat,children,
                           education,employment,wave_id,city_id,both)
  sd_ppl<-distinct(sd_ppl)
  table(sd_ppl$wave_id,sd_ppl$city_id) # one participant from Saskatoon in wave 1 has gone missing?
```
total unique particiapnts across both waves
```{r}
  length(unique(sd_ppl$interact_id))
  table(sd_ppl$wave_id)
```

@@

total unique particiapnts across both waves
```{r}
  length(unique(sd_ppl_valid$interact_id))
```

@@

Plot of stringency index over time (in black and white)
```{r,echo=FALSE}
  si_plot<-ggplot(si[si$province %in% c("British Columbia","Saskatchewan","Quebec") & 
                      si$date <= "2021-02-28",],aes(x=date,y=stringencyIndex*10,color=province)) +
           geom_rect(aes(xmin=ymd('2020-09-10'),
                  xmax = ymd('2021-02-28'),
                  ymin = -Inf,
                  ymax = Inf), fill = 'grey95', color='grey95',alpha = 0.75) +
           geom_line(size=0.6) + 
           theme_classic() + 
           scale_colour_grey(labels=c("British Columbia\n(Vancouver)\n",
                                      "Quebec\n(Montreal)\n",
                                      "Saskatchewan\n(Saskatoon)\n")) +
           labs(#title="Stringency of COVID-19 restrictions over time",
                 x="Date",y="Stringency index") +
           guides(color=guide_legend(title="Province (City)\n")) +
           theme(legend.title=element_text(face="bold",size=9),
                 legend.text = element_text(size=8),
                 axis.title = element_text(face="bold",size=9),
                 axis.text = element_text(size=8),
                 plot.title = element_text(face="bold",size=10,hjust=0.5)) +
           annotate(geom="text", x=as.Date('2020-12-04'), y=10, label="Phase 2 data collection period",
                    color="grey30",size=2.75)

@@

  ggsave(si_plot,file="figures//suppFigure3-20240909.png",width=5.5,height=3.25)
```

@@

**Figure 1**: Distribution of daily minutes of wear time and moderate-to-vigorous physical activity among valid person-days, by wave
```{r,echo=FALSE}
wave.labs<-c("Phase 1","Phase 2")
names(wave.labs)<-c("wave 1","wave 2")

  wear_plot<-ggplot(sd_valid,aes(x=wear)) + 
    geom_histogram(bins=50,aes(fill=wave_id,color=wave_id)) +
    theme_classic() +
    labs(x="Minutes",
         y="Frequency",
         title="Wear time") +
    #geom_vline(xintercept=600,color="black",linetype="dashed") +
    facet_grid(vars(wave_id),
               labeller = labeller(wave_id=wave.labs)) +
    scale_fill_grey(labels=c("Phase 1","Phase 2")) +
    scale_color_grey(labels=c("Phase 1","Phase 2"))+
    theme(axis.title.x = element_text(face="bold",
                                       size=10),
          axis.text.x = element_text(angle=90,vjust = 0.5, hjust=1,size=8),
        # axis.title.y = element_text(#face="bold",
        #                              size=10),
          axis.title.y = element_blank(),
          axis.text.y = element_text(size=8),
          plot.title = element_text(face=2,
                                    hjust=0.5,
                                    size=10),
          legend.position="none",
          strip.text.y = element_text(face=2))

  mvpa_plot<-ggplot(sd_valid,aes(x=mean_mvpa_sd)) + 
    geom_histogram(bins=50,aes(fill=wave_id,color=wave_id)) +
    theme_classic() +
    labs(x="Minutes",
         y="Frequency",
         title="Moderate-to-vigorous physical activity") +
    facet_grid(vars(wave_id),
               labeller = labeller(wave_id=wave.labs)) +
    scale_fill_grey(labels=c("Phase 1","Phase 2")) +
    scale_color_grey(labels=c("Phase 1","Phase 2"))+
    theme(axis.title.x = element_text(face="bold",
                                       size=10),
          axis.text.x = element_text(angle=90,vjust = 0.5, hjust=1,size=8),
        # axis.title.y = element_text(#face="bold",
        #                              size=10),
          axis.title.y = element_blank(),
          axis.text.y = element_text(size=8),
          plot.title = element_text(face=2,
                                    hjust=0.5,
                                    size=10),
          legend.position="none",
          strip.text.y = element_text(face=2))

  titleY=ggpubr::text_grob("Frequency",
                            size =10,rot=90,face="bold")

  suppFigure2<-grid.arrange(wear_plot,mvpa_plot,ncol=2,
               #top=titleMain, # removing title for IJBNPA (titles should be in the text, not in the figure)
               left=titleY)

  ggsave(suppFigure2,file="figures//suppFigure2-20240909.png",width=6,height=4)
```

**Figure 1:** Trend in daily minutes of moderate-to-vigorous physical activity for each wave
```{r,echo=FALSE,warning=FALSE}
  sd_valid$doy3<-sd_valid$date
  sd_valid$doy3<-gsub("2018|2020","2000",sd_valid$doy3)
  sd_valid$doy3<-gsub("2019|2021","2001",sd_valid$doy3)
  sd_valid$doy3<-as.Date(sd_valid$doy3)

  figure1<-ggplot(sd_valid[sd_valid$wave_id=="wave 1",],aes(x=doy3,y=mean_mvpa_sd,color=wave_id)) +
                          geom_point(alpha=0.15,size=1)+
                          geom_point(data=sd_valid[sd_valid$wave_id=="wave 2",],aes(x=doy3,y=mean_mvpa_sd),
                                     alpha=0.15,size=1) +
                          scale_colour_manual(values=c("wave 1"="grey65",
                                                       "wave 2"="black"),
                                              labels=c("  Phase 1 (2018-2019)","  Phase 2 (2020-2021)")) +
                          geom_smooth(method=lm)+
                          geom_smooth(data=sd_valid[sd_valid$wave_id=="wave 2",],aes(x=doy3,y=mean_mvpa_sd),method=lm) +
                          theme_classic()+
                          labs(y = "Daily minutes of moderate-to-vigorous physical activity",
                               x = "Date")+
                          guides(color=guide_legend("")) +
                          theme(axis.title = element_text(face="bold",size=10),
                                axis.text.x = element_text(angle=90,vjust = 0.5, hjust=1,size=8),
                                axis.text.y = element_text(size=8),
                                plot.title = element_text(face="bold",hjust=0.5,size=10),
                                plot.subtitle = element_text(hjust=0.5,size=10),
                                legend.position="bottom",
                                legend.key.size = unit(1,"line"),
                                legend.spacing.x = unit(0.25, 'cm'),
                                legend.margin=margin(-3, 0, 0, 0),
                                legend.text = element_text(size=8,margin = margin(r = 0.4, unit = 'cm'))) 
  figure1
  ggsave(figure1,file="figures//figure1-20240909.png",width=6,height=4.5)
```

@@

*Model: Adjusting for city, age, employment status, income, education, temp, precipitation and wear time*
```{r,warning=FALSE}
  wave_covar_valid_fully<-glmer.nb(mean_mvpa_sd~ wave_id + city_id + wearRS +
                                                 age_cat + employment + income + 
                                                 mean_temp + total_precip + education +
                                                 (1|interact_id),
                          data=sd_valid[sd_valid$age_cat!="missing",])
summary(wave_covar_valid_fully)
```

@@

**Model**
```{r,warning=FALSE}
  both_valid_adj_wear<-glmer.nb(mean_mvpa_sd~ wave_id + 
                                              gender + age_cat + income +
                                              mean_temp + total_precip + wear +
                                       (1|interact_id),
                          data=sd_valid_both) 
summary(both_valid_adj_wear)
```

@@

#### Plot effect modifiers of interest (Figure 2): Income, employment status and education

@@

# Add a income-wave interaction term to the adjusted model  
income_covar_valid_final<-glmer.nb(mean_mvpa_sd~ wave_id + city_id + wave_id*income + wearRS +
                                                age_cat + employment + income + education +
                                                mean_temp + total_precip + 
                                                (1|interact_id),
                          data=sd_valid[sd_valid$age_cat!="missing",])

@@

# create factor variables for both wave_id and income             
income_Inter.HandPick$Income <- factor(income_Inter.HandPick$income,
                      levels=c("<$50,000","$50,000 to $99,999","$100,000 or more","don't know/prefer not to answer"
                               ),
                      labels=c("<$50,000","$50,000-$99,999","$100,000+","No answer"
                               ))
income_Inter.HandPick$wave <- factor(income_Inter.HandPick$wave_id,
              levels=c("wave 1","wave 2"),
              labels=c("Phase 1", "Phase 2"))

@@

# create factor variables for both wave_id and education level          
education_Inter.HandPick$Education <- factor(education_Inter.HandPick$education,
                      levels=c("trade school or college","secondary school","university or graduate degree"),
                      labels=c("College/trade school","High school","University/grad degree"))
education_Inter.HandPick$wave <- factor(education_Inter.HandPick$wave_id,
              levels=c("wave 1","wave 2"),
              labels=c("Phase 1", "Phase 2"))

@@

# create factor variables for both wave_id and employment status             
employment_Inter.HandPick$Employment <- factor(employment_Inter.HandPick$employment,
                      levels=c("employed","unemployed","retired and not working","other"),
                      labels=c("Employed","Unemployed","Retired","Other"))
employment_Inter.HandPick$wave <- factor(employment_Inter.HandPick$wave_id,
              levels=c("wave 1","wave 2"),
              labels=c("Phase 1", "Phase 2"))

@@

**Create Figure 2**
```{r,echo=FALSE,warning=FALSE}
titleY=ggpubr::text_grob("Daily minutes of moderate-to-vigorous\nphysical activity",size =12,rot=90,face="bold")

em_figures<-grid.arrange(income_Plot.HandPick,
                         employment_Plot.HandPick,
                         education_Plot.HandPick,
            left=titleY,
            widths=c(1,1,1),
            heights=c(1),
            nrow=1)
ggsave(em_figures,file="figures//figure2-20240909.png",width=175,height=85,units="mm")
```

@@

**Model**
```{r,warning=FALSE}
  si_all_cov_fully<-glmer.nb(mean_mvpa_sd~ stringencyIndex + city_id + 
                          mean_temp + total_precip + 
                          age_cat + 
                          gender + 
                          children + 
                          employment + income +
                          (1|interact_id),
                          data=sd_valid[sd_valid$wave_id=="wave 2",])
```
